%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Short Sectioned Assignment
% LaTeX Template
% Version 1.0 (5/5/12)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Frits Wenneker (http://www.howtotex.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[paper=a4, fontsize=11pt]{scrartcl} % A4 paper and 11pt font size

\usepackage[T1]{fontenc} % Use 8-bit encoding that has 256 glyphs
\usepackage{fourier} % Use the Adobe Utopia font for the document - comment this line to return to the LaTeX default
\usepackage[english]{babel} % English language/hyphenation
\usepackage{amsmath,amsfonts,amsthm} % Math packages

\usepackage{lipsum} % Used for inserting dummy 'Lorem ipsum' text into the template

\usepackage{sectsty} % Allows customizing section commands
\allsectionsfont{\centering \normalfont\scshape} % Make all sections centered, the default font and small caps

\usepackage{fancyhdr} % Custom headers and footers
\pagestyle{fancyplain} % Makes all pages in the document conform to the custom headers and footers
\fancyhead{} % No page header - if you want one, create it in the same way as the footers below
\fancyfoot[L]{} % Empty left footer
\fancyfoot[C]{} % Empty center footer
\fancyfoot[R]{\thepage} % Page numbering for right footer
\renewcommand{\headrulewidth}{0pt} % Remove header underlines
\renewcommand{\footrulewidth}{0pt} % Remove footer underlines
\setlength{\headheight}{13.6pt} % Customize the height of the header

\numberwithin{equation}{section} % Number equations within sections (i.e. 1.1, 1.2, 2.1, 2.2 instead of 1, 2, 3, 4)
\numberwithin{figure}{section} % Number figures within sections (i.e. 1.1, 1.2, 2.1, 2.2 instead of 1, 2, 3, 4)
\numberwithin{table}{section} % Number tables within sections (i.e. 1.1, 1.2, 2.1, 2.2 instead of 1, 2, 3, 4)

\setlength\parindent{0pt} % Removes all indentation from paragraphs - comment this line for an assignment with lots of text

%----------------------------------------------------------------------------------------
%	TITLE SECTION
%----------------------------------------------------------------------------------------

\newcommand{\horrule}[1]{\rule{\linewidth}{#1}} % Create horizontal rule command with 1 argument of height

\title{	
\normalfont \normalsize 
\horrule{0.5pt} \\[0.4cm] % Thin tp horizontal rule
\huge Charges, Energies and Potentials \\ % The assignment title
\horrule{2pt} \\[0.5cm] % Thick bottom horizontal rule
}

\author{Tobias Kuhn} % Your name

\date{\normalsize\today} % Today's date or a custom date

\begin{document}

\maketitle % Print the title

%----------------------------------------------------------------------------------------
%	PROBLEM 1
%----------------------------------------------------------------------------------------

\section{Problem One}

\begin{align} 
Q_A &= -1[C] & W_A &= 60[J] \\
Q_B &= -2[C] & W_B &= 10[J] 
\end{align}

Two points A and B exist in a vacuum. They have energies and charges as described above.

\subsection{Subproblem c}
\textbf{Problem: How much work in [J] has to be done to get 2.5 trillion electrons from very far away to point A?}
\\

We start our investigation by guessing that calculating the electrostatic charge of $2.5 \cdot 10^{18}$ electrons.

\begin{equation} \label{eq:1}
e^- = -1.6 \cdot 10^{-19} [C]
\end{equation}

Using \ref{eq:1} for the charge of a single electron we let $Q_C$ be the total charge of the 2.5 trillion electrons and calculate that

\begin{align} \label{eq:2}
Q_C &= 2.5 \cdot 10^{18} \cdot e^- \\ &= -0.4 [C]
\end{align}

In order to calculate how much work it takes to move $Q_C$ from infinitely far away to point A,
we need to know how strong the electric field is at any point on the journey. 
Let $r$ be the distance from point A to any point in 3D space and $\epsilon_0$\footnote{$\epsilon_0 = 8.854 \cdot 10 ^ {-12} [m/F] $} the permitivity in empty space (vacuum),
then the electric field $E$ at any distance r from point A is 

\begin{align} \label{eq:3}
E &= \frac{Q_A}{4 \pi \epsilon_0 r^2}
\end{align}

And the force $F$ at any point is

\begin{align} \label{eq:4}
F &= E \cdot Q_C \\
  &= \frac{Q_A}{4 \pi \epsilon_0 r^2} \cdot Q_C
\end{align}

To get the work it takes to move $Q_C$ towards point A from distance $\infty$ to a distance 0 from point A,
we then take the integral of the force from $\infty$ to 0. Note how we introduce the concept of the mathematical limit
to be able to deal with infinities.

\begin{align} \label{eq:5}
W &= - lim_{x\to\infty} \int_x^0{\frac{Q_A Q_C}{4 \pi \epsilon_0 r^2}} \cdot dr \\
  &= - \frac{Q_A Q_C}{4 \pi \epsilon_0 } lim_{x\to\infty} \int_x^0{\frac{1}{r^2}} \cdot dr \\
  &= - \frac{Q_A Q_C}{4 \pi \epsilon_0 } lim_{x\to\infty} \left( \frac{1}{x} - \frac{1}{0} \right) \\
  &= - \frac{Q_A Q_C}{4 \pi \epsilon_0 } \left( 0 - \frac{1}{0} \right)
\end{align}

Now we find ourselves confronted with the infamous divide by 0 problem and it dawns on us, that bringing a charge
$Q_C$ on top on another charge $Q_A$, where the force between the two charges is known to be

\begin{align} \label{eq:6}
F &= \frac{Q_A  Q_C}{4 \pi \epsilon_0 r^2}
\end{align}

will lead to a infinitely large force pushing $Q_C$ away from $Q_A$. After checking that a integral of the form of

\begin{align} \label{eq:7}
\int_\infty^0{\frac{1}{r^2}} \cdot dr \\
\end{align}

indeed is divergant\footnote{ See example 8 here: http://tutorial.math.lamar.edu/Classes/CalcII/ImproperIntegrals.aspx \\
and result here: https://www.wolframalpha.com/input/?i=integral+of+1\%2Fx\%5E2+from+infinity+to+0 

} we give in to the fact that electrons can not exist on top of each other and reinterpret "How much work is necessary to get
2.5 trillion electrons from infinitely far away to point A" as "How much work is necessary to get
2.5 trillion electrons from infinitely far away to very close to point A".

\begin{align} \label{eq:8}
W &= - lim_{x\to\infty} \int_x^{0.000000001[m]}{\frac{Q_A Q_C}{4 \pi \epsilon_0 r^2}} \cdot dr \\
  &= - \frac{Q_A Q_C}{4 \pi \epsilon_0 } lim_{x\to\infty} \int_x^{0.000000001[m]}{\frac{1}{r^2}} \cdot dr \\
  &= - \frac{Q_A Q_C}{4 \pi \epsilon_0 } lim_{x\to\infty} \left( \frac{1}{x} - \frac{1}{0.000000001[m]} \right) \\
  &= - \frac{Q_A Q_C}{4 \pi \epsilon_0 } \left( 0 - \frac{1}{0.000000001[m]} \right) \\
  &= - \frac{-1[C] \cdot -0.4[C]}{4 \pi \epsilon_0 } \left( - \frac{1}{0.000000001[m]} \right) \\
  &= - \frac{-1[C] \cdot -0.4[C]}{4 \pi 8.854 \cdot 10 ^ {-12} [F/m] } \left( - \frac{1}{0.000000001[m]} \right) \\
  &= \frac{0.4 \cdot 10 ^ {12} [C^2]}{4 \pi \cdot 8.854 [F]} \left( \frac{1}{0.000000001} \right) \\
  &= 3.595 \cdot 10^{18} \frac{[C^2]}{[F]} \\
  &= 3.595 \cdot 10^{18} [C \cdot V] \\
  &= 3.595 \cdot 10^{18} \left[A \cdot s \cdot \frac{N \cdot m}{A \cdot s}\right] \\
  &= 3.595 \cdot 10^{18} [N \cdot m] \\
  &= 3.595 \cdot 10^{18} [J]
\end{align}

We come to the conclusion that it takes $3.5950 \cdot 10^{18} [J]$ to get 2.5 trillion electrons from infinitely far away
to 1 [nm] near point A. \\

Note that we only considered the field from $Q_A$ even though we know that there is also a charge
$Q_B$ emitting an electric field $E_B$. We could take field $E_b$ into account if we knew the position in 3D space relative to
point A. But we don't. So we can't. Sounds fishty? Yeah, let's give this another try.


%------------------------------------------------

\section{ Problem One, Second Iteration}
This time the author is confident that he is on point with
how the problem is meant to be solved; Ignoring the electric fields from charges $Q_A$ and $Q_B$.
Which is not to say that there is no electric field at all.
There is one -- because without, the charges wouldn't have potential energies -- but it is implicit. There is no need to calculate it. \\

\subsection{Subproblem c}

\textbf{Problem: How much work in [J] has to be done to get 2.5 trillion electrons from very far away to point A?}
\\

The trick here is to use the conservation of energy. Charge $Q_A$ has a potential
energy of 60[J] because 60[J] of work was done to get it there (pushing it against the implicit static electric field).
\\

We observe that it took 60[J] to get a charge of -1[C] to point A. since the work necessary to move a
charge in a electric field is proportional to the amount of charge, we come to the conclusion
that the work necessary to get a charge of -0.4[C] to point A must be 24[J] (ignoring the field from the charge $Q_A$).

\subsection{Subproblem a}
\textbf{Problem: Calculate the electric voltage between the points.}

\begin{quotation}
According to classical electrostatics, electric potential is a scalar quantity 
denoted by V or occasionally $\phi$ equal to the electric potential energy of any 
charged particle at any location (measured in joules) divided by the charge of that 
particle (measured in coulombs).
\end{quotation}

Above qoute is from the english Wikipedia entry on the electric potential.
To solve this problem we leverage the power of classical electrostatics and proclaim that the
following statement is true.

\begin{align} \label{eq:8}
\phi = \frac{W}{Q}
\end{align}




\end{document}